---
title: "General Social Survey"
author: "Demian Romero"
date: "7 Oct 2019"
output: 
  html_document:
    number_sections: true
    toc: true
---


```{r include=FALSE,echo=FALSE}
require(tigerstats)
require(tidyverse)
require(foreign)
Gss <- read.spss("GSS2018.sav")
Gss1 <- as.tibble(Gss)
```

```{r include=FALSE,echo=FALSE}
#names(Gss1)
mydata <- select(Gss1,VOTE16,RACE,REGION,EDUC,BIGBANG,EVOLVED,DEGREE,POSTLIFE,AGE,OWNGUN,CAPPUN,COURTS,POLVIEWS,VOTE12,GOD)
mydata$AGE <- as.numeric(mydata$AGE)
mydata$EDUC <- as.numeric(mydata$EDUC)
mydata2 <-filter(mydata,RACE!="IAP" ) %>% droplevels()
mydata3 <- filter(mydata2,BIGBANG=="True" | BIGBANG=="False",EVOLVED=="True"|EVOLVED=="False",POSTLIFE=="YES"|POSTLIFE=="NO",OWNGUN=="YES"|OWNGUN=="NO",COURTS=="TOO HARSH"|COURTS=="NOT HARSH ENOUGH"|COURTS=="ABOUT RIGHT") %>% droplevels()
mydata4 <- filter(mydata3,CAPPUN=="FAVOR"|CAPPUN=="OPPOSE",VOTE12=="Voted"|VOTE12=="Did not vote",VOTE16=="Voted"|VOTE16=="Did not vote") %>% droplevels()
levels(mydata4$VOTE12)[1] <- "voted12"
levels(mydata4$VOTE12)[2] <- "no in 12"

```



# Introduction Question 1

<!--In this section you explain what you are trying to show.  Where did the data come from?  What is the research or other question you are trying to answer?!-->

This data is from the 2018 General Social Survey.  Is opinion on the death penalty (CAPPUN) independent of gun ownership (OWNGUN)?


## Methods

<!--Decide on your methods:  use "variable analysis" or other appropriate descriptors.  Make sure to choose at least one graphical method and at least one numerical method.!-->
Both variables in this analysis are categorical; hence we use bar graphs and tables to summarize the data, and a chi-squared test for inference. 

$H_0$ : There is no relationship between gun ownership and whether one is in favor of capitol punishment.

$H_A$ There is a relationship between gun ownership and whether one is in favor of capitol punishment.

##Results

<!--Divide this section into two sub-sections:  One for your descriptive  results and one for your inferential results.!-->

### Descriptive Results

#### Graphical Descriptive Results

<!--Graphical results here.  Make sure to show your code.  Provide appropriate labels for axes, giving units if possible, and provide a good title for the graph, too.  Use the graphical results to describe the patterns if any that exist in the data as focused toward the research question!-->
```{r}
barchartGC(~CAPPUN+OWNGUN, data=mydata4, main='Capitol Punishment and Gun Ownership', type= 'percent')
```

This bar chart shows that those who favor capitol punishment and own guns are about equal, but there are more who do not own a gun than oppose capitol punishment. The bar chart indicates there is a connection between gun ownership and supporting capitol punishment.

#### Numerical Descriptive Results

<!--Numerical results go here. Use the numerical results to describe the patterns if any that exist in the data as focused toward the research question!-->
```{r}
tab1= xtabs(formula = ~CAPPUN+OWNGUN, data=mydata4,)
rowPerc(tab1)
```



### Inferential Results

<!--State hypothesis clearly.  Make sure your discussion of the inferential test covers all the aspects that the test output produces, such as test statistic, p-value etc.  Make a decision about the null hypothesis, explain the assumptions on which the selected test/procedure was based, and why the chosen procedure satisfys the assumptions and is appropriate to answer the research question!-->

```{r}
chisqtestGC(tab1)
```

The p-value is 0.02 < 0.05 so we reject the null hypothesis and conclude we have evidence to support the connection that gun ownership and whether one favors capitol punishment are related. The bar chart indicates there is a connection between gun ownership and supporting capitol punishment.

# Introduction Question 2

<!--In this section you explain what you are trying to show.  Where did the data come from?  What is the research or other question you are trying to answer?!-->

This data is from the 2018 General Social Survey. Now answer Is belief in the big bang theory (BIGBANG) independent of belief in the afterlife (POSTLIFE)

## Methods

<!--Decide on your methods:  use "variable analysis" or other appropriate descriptors.  Make sure to choose at least one graphical method and at least one numerical method.!-->
Both variables in this analysis are categorical; hence we use bar graphs and tables to summarize the data, and a chi-squared test for inference. 

$H_0$ : There is no relationship between belief in the big bang theory and belief in the afterlife.

$H_A$ There is a relationship between belief in the big bang theory and belief in the afterlife.


##Results

<!--Divide this section into two sub-sections:  One for your descriptive  results and one for your inferential results.!-->

### Descriptive Results

#### Graphical Descriptive Results

<!--Graphical results here.  Make sure to show your code.  Provide appropriate labels for axes, giving units if possible, and provide a good title for the graph, too.  Use the graphical results to describe the patterns if any that exist in the data as focused toward the research question!-->

```{r}
barchartGC(~BIGBANG+POSTLIFE, data=mydata4, main='Belief in the Big Bang Theory and belief in the Afterlife ', type= 'percent')
```

The bar chart shows that there does not appear to be a relationship between belief in the Big Bang Theory and belief in the Afterlife since there is little change in each of the data sets.

#### Numerical Descriptive Results

<!--Numerical results go here. Use the numerical results to describe the patterns if any that exist in the data as focused toward the research question!-->

```{r}
tab2= xtabs(formula = ~BIGBANG+POSTLIFE, data=mydata4,)
rowPerc(tab2)
```


### Inferential Results

<!--State hypothesis clearly.  Make sure your discussion of the inferential test covers all the aspects that the test output produces, such as test statistic, p-value etc.  Make a decision about the null hypothesis, explain the assumptions on which the selected test/procedure was based, and why the chosen procedure satisfys the assumptions and is appropriate to answer the research question!-->
```{r}
chisqtestGC(tab2)
```

The p-value is 0.08 > 0.05 so we fail to reject the null hypothesis and conclude we do not have evidence to support the connection. The bar chart does not support a connection between belief in the Big Bang Theory and belief in the Afterlife.


# Introduction Question 3

<!--In this section you explain what you are trying to show.  Where did the data come from?  What is the research or other question you are trying to answer?!-->

This data is from the 2018 General Social Survey. Finally determine if a persons opinion on death penalty (CAPPUN) is independent of their race (RACE).

## Methods

<!--Decide on your methods:  use "variable analysis" or other appropriate descriptors.  Make sure to choose at least one graphical method and at least one numerical method.!-->

Both variables in this analysis are categorical; hence we use bar graphs and tables to summarize the data, and a chi-squared test for inference. 

$H_0$ : There is no relationship between opinion on the death penalty and race.

$H_A$ There is a relationship between opinion on the death penalty and race.

##Results

<!--Divide this section into two sub-sections:  One for your descriptive  results and one for your inferential results.!-->

### Descriptive Results

#### Graphical Descriptive Results

<!--Graphical results here.  Make sure to show your code.  Provide appropriate labels for axes, giving units if possible, and provide a good title for the graph, too.  Use the graphical results to describe the patterns if any that exist in the data as focused toward the research question!-->
```{r}
barchartGC(~CAPPUN+RACE, data=mydata4, main='Belief in capitol punishment by race', type= 'percent')
```


The bar chart shows that there is a connection between belief in capitol punishment and race, since white favors capitol punishment and black opposes capitol punishment. 

#### Numerical Descriptive Results

<!--Numerical results go here. Use the numerical results to describe the patterns if any that exist in the data as focused toward the research question!-->

```{r}
tab3= xtabs(formula = ~CAPPUN+RACE, data=mydata4,)
rowPerc(tab3)
```


### Inferential Results

<!--State hypothesis clearly.  Make sure your discussion of the inferential test covers all the aspects that the test output produces, such as test statistic, p-value etc.  Make a decision about the null hypothesis, explain the assumptions on which the selected test/procedure was based, and why the chosen procedure satisfys the assumptions and is appropriate to answer the research question!-->
```{r}
chisqtestGC(tab3)
```

Since our p-value is .00009 < .05 so we reject the null hypothesis and conclude we have evidence to support the connection between the opinion on capitol punishment and race.

# Discussion and Conclusion

<!--Discussion and conclusion here.  If you found a relationship be sure to consider whether the relationship occurs because one of the variavbles causes the other, or whether they perhasps are related for some other reason.  Watch the chapter 6 videos from the GeorgeTown videos collection.!-->

For question 1: Is opinion on the death penalty independent of gun ownership?: the p-value is 0.02 < 0.05 so we reject the null hypothesis and conclude we have evidence to support the connection that gun ownership and whether one favors capitol punishment are related. The bar chart indicates there is a connection between gun ownership and supporting capitol punishment.

For question 2: Is belief in the big bang theory independent of belief in the afterlife?: The p-value is 0.08 > 0.05 so we fail to reject the null hypothesis and conclude we do not have evidence to support the connection. The bar chart does not support a connection between belief in the Big Bang Theory and belief in the Afterlife.

For question 3: determine if a persons opinion on death penalty is independent of their race: Since our p-value is .00009 < .05 so we reject the null hypothesis and conclude we have evidence to support the connection between the opinion on capitol punishment and race. The bar chart also supports the connection.